---
title: "Cross-Linguistic Comparison of Developmental Trajectories in Vocabulary Composition"
author: "Mika Braginsky, Daniel Yurovsky, Virginia Marchman, and Michael Frank"
date: "2015-05-06"
output:
  html_document:
    highlight: tango
    theme: spacelab
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE, message=FALSE, warning=FALSE)
font = "Open Sans"
```

Load required libraries.
```{r libraries, cache=FALSE}
library(ggplot2)
library(grid)
library(magrittr)
library(dplyr)
library(tidyr)
library(quadprog)
source('data_loading.R')
```

Load in Wordbank common data.
```{r database, cache=FALSE}
wordbank <- connect.to.wordbank("prod")

common.tables <- get.common.tables(wordbank)

items <- get.item.data(common.tables$wordmapping,
                       common.tables$instrumentsmap,
                       common.tables$category) %>%
  filter(type == "word")

instrument.tables <- get.instrument.tables(wordbank, common.tables$instruments)
```

Get vocabulary composition data for all languages.
```{r vocab_comp_fun}
get.vocab.comp <- function(input_language, input_form) {
  
  lang.vocab.items <- filter(items, language == input_language, form == input_form) %>%
    filter(lexical_category %in% c("nouns", "predicates", "function_words")) %>%
    rename(column = item.id) %>%
    mutate(item.id = as.numeric(substr(column, 6, nchar(column))))
  
  lang.instrument.table <- filter(instrument.tables,
                                  language == input_language,
                                  form == input_form)$table[[1]]
  
  lang.vocab.data <- get.instrument.data(lang.instrument.table,
                                         lang.vocab.items$column) %>%
    left_join(select(lang.vocab.items, item.id, lexical_category, item, definition)) %>%
    mutate(value = ifelse(is.na(value), "", value),
           produces = value == "produces",
           understands = value == "produces" | value == "understands")
  
  num.words <- nrow(lang.vocab.items)
  
  lang.vocab.summary <- lang.vocab.data %>%
    group_by(data_id, lexical_category) %>%
    summarise(production.num = sum(produces),
              production.prop = sum(produces) / length(produces),
              comprehension.num = sum(understands),
              comprehension.prop = sum(understands) / length(understands))
  
  lang.vocab.sizes <- lang.vocab.summary %>%
    summarise(production.vocab = sum(production.num) / num.words,
              comprehension.vocab = sum(comprehension.num) / num.words)
  
  lang.vocab.summary %>%
    left_join(lang.vocab.sizes) %>%
    select(-production.num, -comprehension.num) %>%
    mutate(language = input_language,
           form = input_form)
  
  }
```

```{r vocab_comp}
form.vocab.comp <- function(input_form) {
  bind_rows(sapply(unique(filter(instrument.tables, form == input_form)$language),
                   function(lang) get.vocab.comp(lang, input_form),
                   simplify = FALSE)) %>%
    mutate(lexical_category = factor(lexical_category,
                                     levels=c("nouns", "predicates", "function_words"),
                                     labels=c("Nouns", "Predicates", "Function Words"))) %>%
    gather(measure.var, value,
           production.prop, production.vocab,
           comprehension.prop, comprehension.vocab) %>%
    extract(measure.var, c("measure", "var"), "([[:alnum:]]+)\\.([[:alnum:]]+)") %>%
    spread(var, value)
  }

wg.vocab.comp <- form.vocab.comp("WG")
ws.vocab.comp <- form.vocab.comp("WS") %>%
  filter(measure == "production")

vocab.comp <- bind_rows(ws.vocab.comp, wg.vocab.comp)
```

Show sample size of each instrument.
```{r sample_sizes}
sample_sizes <- vocab.comp %>%
  group_by(language, form, measure, lexical_category) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  select(language, form, n) %>%
  distinct()

kable(sample_sizes)
```

Base plot for looking at vocabulary composition.
```{r base_plot}
base_plot <- function(input_form, input_measure) {
  ggplot(filter(vocab.comp, form == input_form, measure == input_measure),
         aes(x = vocab, y = prop, colour = lexical_category)) +
  facet_wrap(~ language) +
  geom_jitter(size = 0.7) +
  geom_abline(slope = 1, intercept = 0, color = "gray", linetype = "dashed") + 
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2),
                     name = "Proportion of Category\n") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2),
                     name = "\nVocabulary Size") +
  scale_color_brewer(palette = "Set1", name = "Lexical Category") +
  theme_bw(base_size = 12) + 
  theme(legend.position = c(0.068, 0.95),
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 9, lineheight = unit(0.8, "char")),
        legend.key.height = unit(0.8, "char"),
        legend.key.width = unit(0.3, "cm"),
        legend.key = element_blank(),
        legend.background = element_rect(fill = "transparent"),
        text = element_text(family = font))
}
```

Plot WS productive vocabulary composition as a function of vocabulary size for each language.
```{r plot_points_ws, fig.width=10, fig.height=10}
base_plot("WS", "production") + geom_jitter(size = 0.7)
```

Plot WG productive vocabulary composition as a function of vocabulary size for each language.
```{r plot_points_wg_prod, fig.width=10, fig.height=10}
base_plot("WG", "production") + geom_jitter(size = 0.7)
```

Plot WG receptive vocabulary composition as a function of vocabulary size for each language.
```{r plot_points_wg_comp, fig.width=10, fig.height=10}
base_plot("WG", "comprehension") + geom_jitter(size = 0.7)
```

Make an S3 class for a clm (constrained linear model), which uses quadratic programming to run a regression on data with a specified formula, subject to the constraint that the coefficients of the regression sum to 1 (in the future could support arbitrary constraints on the coefficients).
```{r clm}
clm <- function(formula, data, ...) {
  M <- model.frame(formula, data)
  y <- M[,1]
  X <- as.matrix(M[,-1])
  s <- solve.QP(t(X) %*% X, t(y) %*% X, matrix(1, nr=ncol(X), nc=1), 1, meq=1)
  class(s) <- "clm"
  s$formula <- formula
  return(s)
  }

# S3 predict method for clm
predict.clm <- function(object, newdata) {
  M <- as.matrix(model.frame(object$formula[-2], newdata))
  s <- object$solution
  p <- (M %*% s)
  rownames(p) <- NULL
  p[,1]
  }

# S3 predictdf method for clm (called by stat_smooth)
predictdf.clm <- function(model, xseq, se, level) {
  pred <- predict(model, newdata = data.frame(x = xseq))
  data.frame(x = xseq, y = as.vector(pred))
  }
```

Plot WS productive vocabulary composition as a function of vocabulary size for each language with cubic contrained lm curves.
```{r plot_models_ws, fig.width=10, fig.height=10}
base_plot("WS", "production") +
  geom_smooth(method = "clm", formula = y ~ I(x^3) + I(x^2) + x - 1)
```

Plot WG productive vocabulary composition as a function of vocabulary size for each language with cubic contrained lm curves.
```{r plot_models_wg_prod, fig.width=10, fig.height=10}
base_plot("WG", "production") +
  geom_smooth(method = "clm", formula = y ~ I(x^3) + I(x^2) + x - 1)
```

Plot WG receptive vocabulary composition as a function of vocabulary size for each language with cubic contrained lm curves.
```{r plot_models_wg_comp, fig.width=10, fig.height=10}
base_plot("WG", "comprehension") +
  geom_smooth(method = "clm", formula = y ~ I(x^3) + I(x^2) + x - 1)
```

Function for resampling data.
```{r resample}
sample.areas <- function(d, num.times=1000) {
  
  poly.area <- function(group.data) {
    model = clm(prop ~ I(vocab^3) + I(vocab^2) + vocab - 1,
                data = group.data)
    return((model$solution %*% c(1/4,1/3,1/2) - 0.5)[1])
  }
  
  counter = 1
  sample.area <- function(d) {
    d.frame <- d %>%
      group_by(language, form, measure) %>%
      sample_frac(replace = TRUE) %>% # resample kids
      group_by(language, form, measure, lexical_category) %>%
      do(area = poly.area(.)) %>%
      mutate(area = area[1]) %>%
      rename_(.dots = setNames("area", counter))

    counter <<- counter + 1 # increment counter outside scope
    return(d.frame)
  }

  areas <- replicate(num.times, sample.area(d), simplify=FALSE)
  
  Reduce(left_join, areas) %>%
    gather(sample, area, -language, -form, -measure, -lexical_category)
}
```

Resample data, compute area for each sample, find the mean and CI of the area estimate.
```{r areas}
areas <- sample.areas(vocab.comp, 100)

area.summary <- areas %>%
  group_by(language, form, measure, lexical_category) %>%
  summarise(mean =  mean(area),
            ci.high = quantile(area, 0.975),
            ci.low = quantile(area, 0.025)) %>%
  ungroup() %>%
  mutate(language = factor(language))

#nouns <- filter(area.summary, lexical_category == "Nouns")
#noun.levels <- nouns$language[order(nouns$mean, nouns$language, decreasing = FALSE)]
#area.summary %<>% ungroup() %>% mutate(language = factor(language, levels = noun.levels))
```

Plot each lexical category's area estimate by language, form, and measure.
```{r plot_areas, fig.width=12, fig.height=8}
ggplot(area.summary, aes(y = language, x = mean, col = lexical_category)) +
  facet_grid(lexical_category ~ form + measure) +
  geom_point() +
  geom_segment(aes(x = ci.low, xend = ci.high,
                   y = language, yend = language)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") + 
  scale_colour_brewer(palette = "Set1", name = "", guide=FALSE) +
  scale_y_discrete(limits = rev(levels(area.summary$language))) +
  theme_bw() +
  theme(text = element_text(family = font)) + 
  ylab("") +
  xlab("\nRelative representation in early vocabulary") 
```