---
title: About Us
output:
  html_document:
    highlight: tango
    theme: null
---

### What is Wordbank?

Wordbank is an open database of children’s vocabulary growth, featuring data from [contributors around the world](http://wordbank.stanford.edu/contributors). 

Wordbank archives data from the [MacArthur-Bates Communicative Development Inventory (MB-CDI)](http://mb-cdi.stanford.edu), a family of parent-report questionnaires and enables researchers to browse these data in [interactive analyses](http://wordbank.stanford.edu/analyses) and access them via the [wordbankr](http://github.com/langcog/wordbankr) R package. 

### Who funds Wordbank?

The team would like to thank our supporters!  Wordbank is currently funded by a grant from the National Science Foundation (# 1451577), as well as generous support from the MB-CDI Advisory Board.

### Meet the Wordbank Team!


```{r, echo=FALSE}
library(htmltools)

thumbnail <- function(first_name, last_name, institution, img, href) {
  a <- div(class = "col-md-3",
      div(class = "thumbnail",
        if (href != "") {
          a(name = paste0(first_name,"_",last_name), href = href, target="_blank",
            img(src = img)
          )
        } else {
          img(src = img)
        },
        div(class = "caption",
          h3(first_name, br(),last_name),
          p(institution)
        )
      )
  )
  HTML(gsub("&#39;","'", as.character(a)))
}

```

```{r, echo=FALSE}
HTML("{% load staticfiles %}")
div(class="row",
  div(class="col-md-1"),
  
  thumbnail("Michael C.","Frank", "Stanford University", "{% static 'images/headshots/mike.jpg' %}", "http://web.stanford.edu/~mcfrank"),
  
  thumbnail("Mika", "Braginsky", "MIT", "{% static 'images/headshots/mika.jpg' %}", "http://mikabr.github.io"),
  
  thumbnail("Daniel", "Yurovsky", "University of Chicago", "{% static 'images/headshots/dan.jpg' %}", "http://www.thoughtfill.com"),
  
  div(class="col-md-1")
)
br()

div(class="row",
  div(class="col-md-2"),
  
  thumbnail("Virginia","Marchman", "Stanford University", "{% static 'images/headshots/virginia.jpg' %}", "https://web.stanford.edu/group/langlearninglab/cgi-bin/research-associates/Virginia_Marchman,_Ph.D.php"),
  
  thumbnail("Danielle","Kellier", "Stanford University", "{% static 'images/headshots/danielle.jpg' %}", ""),
  
  div(class="col-md-3")
)



```
